Robotic cloth manipulation for clothing assistance task using Dynamic Movement Primitives

### 2017-01-26 paper/paper.tex
[major] add 1st draft

\usepackage{booktabs} % For formal tables

@@

\begin{abstract}
	The need of robotic clothing assistance in the field of assistive robotics is growing, as it is one of the most basic activities in daily life of elderly and disabled people. In this study, we are investigating the applicability of using Dynamic Movement Primitives (DMP) as a task parameterization model for performing clothing assistance tasks. The robotic cloth manipulation task deals with putting the cloth on both the arms. The robot should do cooperative manipulation by holding the cloth. Also, there can be many failure scenarios as clothes are highly non-rigid. DMP can represent nonlinear motion with a set of differential equations. These equations can be adapted to generate any movement trajectory just by changing the goal parameter. The system consists of Baxter humanoid robot and Microsoft Kinect RGBD sensor for tracking the posture of hands. To perform the task, a demonstration is recorded by moving the Baxter arms in the appropriate trajectory. The recorded trajectory is parameterized by using DMP. Once the system is trained, new postures are accommodated by DMP. The cloth manipulation is done by Baxter humanoid robot which follows the trajectory generated by DMP. We have performed the experiments on soft mannequin instead of human. The result shows that DMPs are able to generalize the movement trajectory for the modified posture also.
\end{abstract}

@@

\keywords{Robotic Clothing Assistance, Dynamic Movement Primitives (DMP)}

@@

\section{Introduction}

The \textit{proceedings} are the records of a conference\footnote{This
is a footnote}.  ACM seeks
to give these conference by-products a uniform, high-quality
appearance.  To do this, ACM has some rigid requirements for the
format of the proceedings documents: there is a specified format
(balanced double columns), a specified set of fonts (Arial or
Helvetica and Times Roman) in certain specified sizes, a specified
live area, centered on the page, specified size of margins, specified
column width and gutter size.

\section{Related Works}
Typically, the body of a paper is organized into a hierarchical
structure, with numbered or unnumbered headings for sections,
subsections, sub-subsections, and even smaller sections.  The command
\texttt{{\char'134}section} that precedes this paragraph is part of
such a hierarchy.\footnote{This is a footnote.} \LaTeX\ handles the
numbering and placement of these headings for you, when you use the
appropriate heading commands around the titles of the headings.  If
you want a sub-subsection or smaller part to be unnumbered in your
output, simply append an asterisk to the command name.  Examples of
both numbered and unnumbered headings will appear throughout the
balance of this sample document.

\section{Dynamic Movement Primitives}
Dynamic Movement Primitives (DMP) aims at designing contoller for learning and generalization of Motor Skills by learning from demonstration \cite{ijspeert2013dynamical}. The controllers are based on nonlinear dynamical systems, and use locally weighted regression techniques to learn complex, discrete or rhythmic, movements demonstrated by a human subject. These controllers can be considered to be discrete or rhythmic pattern generators which can replay and modulate the learned movements, while being robust against perturbations.

@@

This way the DMP can be made to imitate the desired path.

\section{Overview of the System}
Robotic cloth manipulation task contains a dual arm humanoid robot Baxter. The complete setup is shown in figure \ref{setup}. We choose soft mannequin instead of a human for this preliminary experiment. Both the arms of mannequin are open and  given the support by a metallic stand, to avoid falling down the arms. The mannequin is positioned in such a way, so that it resides within the limits of work space of the Baxter robot. Also both the arms of mannequin are facing towards robot. A kinect v2 sensor is mounted on the LCD display of Baxter root. Kinect sensor can see the mannequin and clothing article. Before starting the experiment, the clothing article is put in arms of the baxter robot manually.

The Baxter robot is connected to a computer directly using Ethernet cable. It is controlled using Robot Operating System (ROS), one of the widely used tool by the researchers in robotics community. We used Baxter robot's API, which are available and supported by ROS \cite{fitzgerald2013developing} to command the robot. The Kinect sensor is also controlled by Open source Kinect API for ROS.

@@

\section{Experiments}
As per the formulation \ref{setup}, the DMP can learn by the demonstration. Hence we starts by performing a demonstration by holding the robot arm and move accordingly. During the demonstration, pose of the end-effector is recorded. The term pose collectively refers to position and orientation. Once the demonstration is finished, DMP is initialized using the recorded trajectory. Three DMP trajectories one for each coordinate axis are initialized for one arm. In this way, we have totally six DMP trajectories, which can control both the arms of Baxter robot. We performed following two experiments by using these trajectory- (a) Clothing task using position DMP (b) Failure detection using end-effector forces.

\subsection{Clothing task using position DMP}
The aim of this experiment is to put the clothing article on both the arms of mannequin by using DMP system. We use the position data to initialize the DMP trajectories, which are being used in this task. The posture of mannequin is changed by lifting the arms up or down. At this point, we use Kinect Sensor to get the 3D coordinates of the arm. Now we change the goal of DMP trajectories by using this information. The modified DMP can be acquired from equation 1.

@@

\section{Results}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.

\begin{acks}
	The authors would like to thank Dr. Yuhua Li for providing the
	matlab code of  the \textit{BEPS} method.

	The authors would also like to thank the anonymous referees for
	their valuable comments and helpful suggestions. The work is
	supported by the \grantsponsor{GS501100001809}{National Natural
		Science Foundation of
		China}{http://dx.doi.org/10.13039/501100001809} under Grant
	No.:~\grantnum{GS501100001809}{61273304}
	and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young
	Scientsts' Support Program}.

\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sigproc}

>>>>

%\usepackage{booktabs} % For formal tables
\usepackage{subcaption}

@@

\begin{abstract}
	The need of robotic clothing assistance in the field of assistive robotics is growing, as it is one of the most basic activities in daily life of elderly and disabled people. In this study, we are investigating the applicability of using Dynamic Movement Primitives (DMP) as a task parameterization model for performing clothing assistance tasks. The robotic cloth manipulation task deals with putting the cloth on both the arms. The robot should do cooperative manipulation by holding the cloth. Also, there can be many failure scenarios as clothes are highly non-rigid. DMP can represent nonlinear motion with a set of differential equations. These equations can be adapted to generate any movement trajectory just by changing the goal parameter. The system consists of Baxter humanoid robot and Microsoft Kinect RGBD sensor for tracking the posture of hands. To perform the task, a demonstration is recorded by moving the Baxter arms in the appropriate trajectory. The recorded trajectory is parameterized by using DMP. Once the system is trained, new postures are accommodated by DMP. The cloth manipulation is done by Baxter humanoid robot which follows the trajectory generated by DMP. We have performed the experiments on soft mannequin instead of human. The result shows that DMPs are able to generalize the movement trajectory for the modified posture.
\end{abstract}

@@

\keywords{Robotic Clothing Assistance, Dynamic Movement Primitives (DMP), Human-Robot Interaction, Learning and Adaptive Systems}

@@

\section{Introduction}
\label{introduction}
Due to the demographic trend in developed countries, the robotic assistance in the filed of elderly care is in home environments growing \cite{broekens2009assistive}. Although there has been a significant number of research done in this field, robotic clothing assistance, one of the basic and important assistance activities in daily life of elderly as well as disabled people is yet an open field for research. While rigid object manipulation with robots has mainly relied on precise robot control, the deformable objects are rather require a complex control scheme. The clothing assistance is a challenging problem since robots must interact with non-rigid and highly deformable clothes, and with the assisted person whose posture can vary during the assistance.

In this study, we are investigating the applicability of using Dynamic Movement Primitives (DMP) as a task parameterization model for performing clothing assistance tasks. The robotic cloth manipulation task deals with putting the cloth on both the arms. The idea of using DMP is inspired from the fact that DMP can learn the complex task from the demonstration \cite{ijspeert2003learning, schaal2006dynamic, ijspeert2013dynamical} and thus reduce the manual efforts to design the controller from scratch or to fine-tune the controller parameters. We choose the dual arm Baxter robot in this research as it is safe and flexible by the design \cite{fitzgerald2013developing}.

The rest of the paper is organized as follows. Section \ref{related_works}, gives a brief overview about the related literature in this field. Section \ref{dmp} introduces the mathematical formulation about Dynamic Movement Primitives. In Section \ref{system_overview}, we describe our system and various components used in the experiments. Section \ref{experiments} deals with the details about experiments performed. Section \ref{results} shows the experimental results. Finally we conclude in Section \ref{conclusions} with some future directions.

\section{Related Works}
\label{related_works}
In this section, we provide brief overview of the related literature in this field. There has been significant number of research done in the field of Robotic Clothing Assistance. Colom{\'e} et al. \cite{colome2015friction} proposed a framework for Reinforcement Learning of Robotic Tasks in non-rigid environments. They performed the task of wrapping a scarf around the neck of a mannequin and used color based segmentation to distinguish mannequin, scarf and a mark placed on the nose of mannequin. The main focus of their work is to incorporate friction based model while performing clothing task. Gao et al. \cite{gao2015user, gao2016iterative} has focused on user upper-body modeling for personalized dressing by using top-view depth camera. Randomised decision forests was used to estimate user pose and proposed an online iterative path optimisation method to enable Baxter humanoid robot to assist human to wear a sleeveless jacket. Another interesting work by Kapusta et al. \cite{kapusta2016data} is focused towards designing a controller inspired from data-driven haptic perception. They classified the forces measured at robot's end effector by using hidden Markov models and performed the clothing task on hospital gown. Their focus was to classify force data for haptic perception with high accuracy. Klee et al. \cite{klee2015personalized} worked on personalized assistance for dressing a user, where they used vision module to monitor the human's motion. The robot request to user to move towards robot, monitors motion and puts hat on the user once users is rechable by the robot. Koganti et al. \cite{koganti2015cloth} proposed a framework for offline learning of cloth dynamics model using Gaussian Process Latent Variable Models (GP-LVM) by incorporating motion capture data and applying this model for the online tracking of human-cloth relationship using a depth sensor. They showed that the shared GP-LVM is able to learn reliable motion models of the T-shirt state for robotic clothing assistance tasks. Representing cloth state in low-dimensional field by using topology coordinates is another impressive work done by Tamei et al. \cite{tamei2011reinforcement}. They proposed Reinforcement Learning framework and demonstrated that the robot quickly learns a suitable arm motion for putting T-shirt into the mannequin's head. Yamazaki et al. \cite{yamazaki2013method, yamazaki2014bottom}  worked on bottom dressing by a Life-sized humanoid robot, where they recognize the cloth state by using  optical flow on the images acquired from single camera. They showed that the robot can pull a bottom clothing item along the subject's legs.

Another related filed of research is Motor-skills Learning for Cloth Handling, in which Yamakawa et al. \cite{yamakawa2011dynamic} proposed a new strategy for dynamic manipulation of sheet-like flexible objects by a high-speed robot system. The system consists of two high-speed multi-fingered hands mounted on two sliders and a high-speed vision system. The proposed system learns the necessary motor skills from the demonstration performed by a human subject. They validated the robot trajectory obtained from the motion planning method was with simulation results. Another exciting work done by Mons{\'o} et al. \cite{monso2012pomdp}, where they proposed a probabilistic planner, based on Partially Observable Markov Decision Process (POMDP) approach, for reducing the inherent uncertainty of cloth sorting (isolation/extraction) task. Their approach relaxes the precision requirements of robot vision and manipulation.

\section{Dynamic Movement Primitives}
\label{dmp}
Dynamic Movement Primitives (DMP) aims at designing contoller for learning and generalization of Motor Skills by learning from demonstration \cite{ijspeert2003learning}. The controllers are based on nonlinear dynamical systems, and use locally weighted regression techniques to learn complex, discrete or rhythmic, movements demonstrated by a human subject \cite{ijspeert2002movement}. These controllers can be considered to be discrete or rhythmic pattern generators which can replay and modulate the learned movements, while being robust against perturbations.

@@

This way the DMP can be made to imitate the desired path. However, this formulation of DMP suffers from critical issues when used in practical application, such as when the scaling term of $f$, $g-x_0$ if close to zero, a small change in $g$ may lead to huge acceleration and thus it may break the limits of robot. We used the improved version of DMP \cite{pastor2009learning} in this research.

\section{Overview of the System}
\label{system_overview}
Robotic cloth manipulation task contains a dual arm humanoid robot Baxter. The complete setup is shown in figure \ref{setup}. We choose soft mannequin instead of a human for this preliminary experiment. Both the arms of mannequin are open and  given the support by a metallic stand, to avoid falling down the arms. The mannequin is positioned in such a way, so that it resides within the limits of work space of the Baxter robot. Also both the arms of mannequin are facing towards robot. A kinect v2 sensor is mounted on the LCD display of Baxter root. Kinect sensor can see the mannequin and clothing article. Before starting the experiment, the clothing article is put in arms of the baxter robot manually.

The Baxter robot is connected to a computer directly using Ethernet cable. It is controlled using Robot Operating System (ROS), one of the widely used tool by the researchers in robotics community. We used Baxter robot's API, which are available and supported by ROS to command the robot. The Kinect sensor is also controlled by Open source Kinect API for ROS.

@@

\section{Experiments}
\label{experiments}
As per the formulation \ref{setup}, the DMP can learn by the demonstration. Hence we starts by performing a demonstration by holding the robot arm and move accordingly. During the demonstration, pose of the end-effector is recorded. The term pose collectively refers to position and orientation. Once the demonstration is finished, DMP is initialized using the recorded trajectory. Three DMP trajectories one for each coordinate axis are initialized for one arm. In this way, we have totally six DMP trajectories, which can control both the arms of Baxter robot. We performed following two experiments by using these trajectory- (a) Clothing task using position DMP (b) Failure detection using end-effector forces.

\subsection{Clothing task using position DMP}
The aim of this experiment is to put the clothing article on both the arms of mannequin by using DMP system. We use the position data to initialize the DMP trajectories, which are being used in this task. The posture of mannequin is changed by lifting the arms up or down. At this point, we use Kinect Sensor to get the 3D coordinates of the arm. Now we change the goal of DMP trajectories by using this information. The modified DMP can be acquired as described in section \ref{dmp}.

@@

\section{Results}
\label{results}
A soft mannequin was used in order to perform the clothing task. A sleeveless tee shirt was used during the experiment and the task was to put the clothing article on both the arms as shown in figure \ref{various_posture}. Following subsections explains the results of the two experiments described above.

\begin{figure}
	\includegraphics[width=0.45\textwidth]{various_posture}
	\caption{Various postures used in Clothing task}
	\label{various_posture}
\end{figure}

\begin{figure}
	\includegraphics[width=0.45\textwidth]{trajectory}
	\caption{Left arm trajectory of Baxter}
	\label{trajectory}
\end{figure}

\subsection{Clothing task using position DMP}
In this experiment, the initialized DMP was modified to accommodate the new posture by changing the goal state of the DMP. The generated trajectory was then run on the Baxter robot as shown in figure \ref{trajectory}. The newly generated DMP trajectory (shown in red color) was not only found well suited and capable of performing clothing task but also smooth compared to the original trajectory (shown in blue color). A video demonstartion of this experiment can be seen at YouTube\footnote{\url{http://youtu.be/Rb2JePazJjk}}.

To eventuate this experiment, we performed it many times for various postures by keeping the arms at different-different height. During the experiment, we monitored the trajectory generated by DMP system. The accuracy measurement is shown in figure \ref{accuracy}.

\begin{figure}
	\centering
	\begin{subfigure}{.2\textwidth}
		\includegraphics[width=\linewidth]{inclination}
	\end{subfigure}%
	\begin{subfigure}{.26\textwidth}
		\includegraphics[width=\linewidth]{accuracy}
	\end{subfigure}
	\caption{Accuracy measurement}
	\label{accuracy}
\end{figure}

\subsection{Failure detection using end-effector forces}
The clothing task has to deal with complex dynamics including manipulation of clothing article. The clothes are non-rigid, flexible and highly deformable objects, making the task more difficult to perform. During the clothing task, we observed the forces being applied at both the end-effector of Baxter robot. The trajectories were monitored and categorized into two \textit{success} and \textit{failure}. The mean of these two category is calculated and plotted as shown in figure \ref{position_force}.

It is clearly visible from the figure \ref{position_force} that the applied forces are very different in nature in both the cases. Both of these applied forces are increasing from the beginning however the applied forces in case of \textit{failure} are much more higher than that of \textit{success}. Hence one can easily differentiate and detect the failure by using this information.

\begin{figure}
	\includegraphics[width=0.45\textwidth]{position_force}
	\caption{Failure detection using end-effector forces}
	\label{position_force}
\end{figure}

\section{Conclusions}
\label{conclusions}
This paper presents an approach for robotic cloth manipulation for clothing assistance task using Dynamic Movement Primitives. A dual arm Baxter robot, soft mannequin and very thin sleeveless tee shirt were used in the task. The authors have also presented an approach for failure detection using forces being applied on the end-effector of the robot. Authors have used the Baxter APIs in order to get the forces, which are calculated by Baxter Dynamics Module. Though raw forces were very noisy in nature, but after applying median filter most of the noise was eliminated properly.

We plan to extend our research to make the approach more robust by adding visual information and force information with DMP system in the future. Also there is a need of designing an adaptive controller for real time tracking to adapt and detect various failure scenarios. Therefore, in the future, a combination of robot vision and force data can provide the better estimation of cloth state.

\begin{acks}
	The authors would also like to thank the anonymous referees for their valuable comments and helpful suggestions. The work is supported by the \grantsponsor{GS501100001809}{National Natural Science Foundation of China}{http://dx.doi.org/10.13039/501100001809} under Grant No.:~\grantnum{GS501100001809}{61273304} and~\grantnum[http://www.nnsf.cn/youngscientsts]{GS501100001809}{Young Scientsts' Support Program}.
\end{acks}

\bibliographystyle{ACM-Reference-Format}
%\bibliographystyle{unsrt}
\bibliography{air}
